---
title: 'Developing Data Products : Course Project'
author: "Bhanu Poudyal"
date: "September 27, 2015"
output: html_document
runtime: shiny
---
##Synpsis
This study examines whether automatic transmission or manual transmission cars are better in terms of mileage per gallon. It also quantify MPG difference between automatic and manual transmission cars. 

###Data Analysis

####Part 1: Is an automatic or manual transmission better for MPG

Smmary of dataset
```{r, echo=TRUE, eval=TRUE}
summary(mtcars)
```

Required variables to get general idea regarding their relation

I could plot the data based on user input. But that action needs reactive expression environment, which slows down the app.  The commented out part could do the job. Hence I used user input just to show how it can be done. The plot is normal qplot of hp vs mpg.
```{r, echo=TRUE, eval=TRUE}
data(mtcars)
inputPanel(
  selectInput("mtCar_VAriables1", label = "First Parameter to Plot:",
              choices = names(mtcars), selected = 20),
 selectInput("mtCar_VAriables2", label = "Second Paraneter to plot",
              choices = names(mtcars), selected = 20) 
)
#renderPlot(input$mtCar_VAriables1, input$mtCar_VAriables2, data=mtcars)
library(ggplot2)
plotOutput("plot")
```

As the dataset we are interested in as binary outcome or dependent variable, it normal linear model may not yield desired result. Hence, we have to use generalized linear model

```{r, echo=TRUE, eval=TRUE}
fit<-glm(am~mpg, data = mtcars, family = "binomial")
coef(fit)
```

From the coefficient above, we can estimate that every unit miles per gallon, the log odds of automatic transmission versus standard transmission increase by 0.307. 
When we employ wald.test as demonstrated in script below, the Chi-squared test statistic of 7.9 with one degree of freedom for p-value 0.005 is statistically significant.

```{r, echo=TRUE, eval=TRUE}
library(aod)
wald.test(vcov(fit), coef(fit), Terms = 1)
```

Identifying odd ratio in 95% confidence interval gives us that we can estimate that every unit miles per gallon the odds of automatic transmission versus standard transmission increase by 1.3594

```{r, echo=TRUE, eval=TRUE}
exp(cbind(OR = coef(fit), confint(fit)))
```

#### Part 2: Quantify the MPG difference between automatic and manual transmissions

Let us list required variables only from mtcars dataset.

```{r, echo=TRUE, eval=TRUE, results="hide"}
library(sqldf)
autoAllTrans<-sqldf("select am, mpg from mtcars")
autoAllTrans
```

Applying the increase in miles to compare manual transmission versus automatic transmission, manual mpg, equivalent automatic mpg and the difference are calculated below:

```{r, echo=TRUE, eval=TRUE}
autoTransVsManual<-sqldf("select am, mpg as Manual,  mpg*1.3594 as EquivalentAutomatic, mpg* 1.3594-mpg as Difference from mtcars where am=0")
autoTransVsManual
```

##Results

Manual cars are better than automatic ones based an MPG. With odd ratio in 95% confidence interval gives us that we can estimate that every unit miles per gallon the odds of automatic transmission versus standard transmission increase by 1.3594.



*This is shiny version of my previous project.