---
title       : Shiny and Slidify- Tools to Present Data Products
subtitle    : Tools at a Glance
author      : Bhanu Pouduyal
job         : Coursera- Developing Data Products
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---

## Shiny and Slidify- Tools to Present Data Products

### Shiny and Slidify work together with R to make data presentation on the worldwide web easy.

--- .class #id 

## Tools at a Glance

Shiny and Slidify make it easy to convert R codes and R markdown documents to present on the Worldwide Web.

* Shiny helps us create simple but interactive web pages that include R codes, their evaulations and HTML document markers in easy way
* Slidify helps us create PowerPoint like presentation with R codes and their evaluation results in very easy way.

--- .class #id 

## Writing and evaluating codes is easy with Slidify

```{r echo=TRUE, eval=TRUE}
data(mtcars)
head(mtcars)
```
This slide shows top few records of mtcars dataset.

--- .class #id 

## We Can Plot Graphs Using Slidify

```{r, echo=TRUE, eval=TRUE}
plot(mtcars$disp, mtcars$mpg, col=c(mtcars$disp, mtcars$mpg), 
     xlab = "Displacement", ylab = "Mileage per gallon")
```

--- .class #id 

## Thank You!

The purpose of this presentation was to show you basic characterististics of Slidiy and how can it be integregated with R code.
Thank You for viewing my slides


